### 2024-03-16 paper/paper.qmd
new updatessss

---
title: "Beyond Capacity: Analyzing Shelter Use and Government Responses to Homelessness in Toronto"
author: 
  - Jerry Lu (Yen-Chia Lu)
  - Sinan Ma 
  - Che-Yu Wang
thanks: "Code and data are available at: https://github.com/Sinanma/shelter_in_Toronto."
date: 12 March, 2024
date-format: long
abstract: "The study examines the impact of the COVID-19 pandemic and government policies on Toronto's shelter system in 2022. It analyzes the system's response to external pressures using a dataset that includes shelter occupancy rates, service user counts, and capacity metrics. The findings show significant fluctuations in shelter demand and occupancy, highlighting both the system's resilience and vulnerabilities. The study also offers policymakers valuable insights into how to strengthen support for homeless populations in the face of future challenges. Despite temporary emergency shelters and a plan to transition to permanent affordable housing, the shelter system was under pressure from the pandemic, opioid crisis, and an increase in refugee claimants. The analysis offers useful insights for policymakers looking to improve support for homeless people."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
pdf-engine: xelatex
---

```{r}
#| include: false
#| warning: false
#| message: false

####workspcae setup####

library(tidyverse)
library(here)
library(rstanarm)
library(modelsummary)
library(knitr)
library(dplyr)
library(tibble)
```

```{r}
#| include: false
#| warning: false

#### import data ####
toronto_shelters_clean <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

# Introduction

During the COVID-19 pandemic, urban shelters faced unprecedented challenges. The pandemic's ripple effects have reached Toronto's shelters, which provide a safety net for the city's homeless population. As public health directives evolved and governmental policies changed, the shelter system's response became a critical area of research to ensure the homeless's safety and well-being during and after the crisis.
This paper looks at the complexities of shelter use, service demand, and capacity in Toronto's diverse urban landscape in 2022. Despite previous research on homelessness, there are few comprehensive analyses that combine a Bayesian approach with a visual examination of service patterns across shelter types and demographics. This study seeks to close that gap by providing a nuanced analysis of the current state of the shelter system using sophisticated statistical models and extensive visual data. On the other hand, we discovered some secondary resources about the city of Toronto, as well as any other pertinent resources. According to a City of Toronto article, the city plans to build over 400 affordable homes for homeless people, with 1,100 supportive homes completed in 2020. By 2022, the City of Toronto expects nearly 3,000 new housing opportunities.However, we also cite the article "Prevalence of SARS-CoV-2 infection among people experiencing homelessness in Toronto during the first wave of the COVID-19 pandemic" by the author Linh Luong[@luong2022prevalence] and others that The COVID-19 pandemic has had a significant impact on homelessness, especially among Indigenous and racialized communities. Over 35,000 people in Canada are homeless on a daily basis, with Toronto's homeless population being especially diverse.

Furthermore, our Bayesian analysis reveals an unexpected pattern: men account for a disproportionately high proportion of all shelter occupants. This is especially noticeable in the colder winter months, indicating a seasonal susceptibility that corresponds to Toronto's lower temperatures. Furthermore, the data show a troubling pattern of men repeatedly using short-term shelters, indicating potential gaps in services designed to promote long-term housing stability.

These discoveries are undeniably significant. The high proportion of men in shelter occupancy data necessitates a thorough reassessment of existing support structures designed specifically for men. It requires that resource allocation and policy making be gender-sensitive. The high turnover rates and seasonal peaks in occupancy among men highlight the need for targeted interventions that address not only immediate shelter needs but also more general systemic issues such as employment, mental health, and substance abuse—all of which are frequently associated with homelessness. By highlighting these issues, our research contributes to a more informed conversation, which is critical for developing strong, long-term solutions for Toronto's homeless men.
# Data {#sec-data}

The data set was collected by the city of Toronto's data @opendatatoronto. We have choose the daily shelter and overnight service occupancy and capacity as the main direction of research. We analysis the data in R(@citeR), with additional tools for support my analysis, including `tidyverse`[@citetidyverse], `here`[@citehere], `dplyr`[@citedplyr], `readr`[@citereadr], `modelsummary`[@citemodelsummary], `janitor`[@citejanitor], `tibble`[@citetibble], `ggplot2`[@citeggplot2], and research method is from @tellingstorieswithdata.

# Average rooms usage in Toronto Shelter of 2022
The statistics from @tbl-torontoo and scatter plot in @fig-torontoo illustrated the mean daily occupancy rates for Toronto's shelter system in 2022, offering a detailed perspective on the impact of seasonal variations and other factors on shelter utilization. The occupancy rate starts at a reasonable level during the winter months and gradually increases, reaching its highest point in October at 71.5, before slightly declining towards the end of the year. This pattern indicates a growing demand for shelter as the year advances, especially during the autumn season. This increase could be linked to cooler temperatures, changes in socioeconomic conditions, or alterations in policies that impact the homeless community.
The following abstract highlights the influence of COVID-19, the opined crisis, and increasing refugee claims on the shelter system, suggesting that these issues are major contributors to the observed changes in shelter usage. Although there have been efforts to make housing more accessible and provide temporary emergency shelters, the shelter system is still being challenged, as shown by the consistently high occupancy rates, especially towards the end of the year. This suggests that the shelter system is both adaptable and under significant strain, highlighting the urgent requirement for policymakers to modify and strengthen their assistance for the homeless population. This is particularly important in anticipation of future issues that could further burden the system.

```{r}
#| label: tbl-torontoo
#| tbl-cap: "Shelter usage in Toronto in 2022"
#| echo: false
  toronto_shelters_clean |>
  mutate(occupancy_month = month(
    occupancy_date,
    label = TRUE,
    abbr = FALSE
  )) |>
  arrange(month(occupancy_date)) |> 
  drop_na(occupied_rooms) |>
  summarise(number_occupied = mean(occupied_rooms),
            .by = occupancy_month) |>
  kable(
    col.names = c("Month", "Avg daily occupied rooms"),
    digits = 1
  )
```

```{r}
#| label: fig-torontoo
#| fig-cap: "Toronto shelter rooms average usage in 2022"
#| echo: false

toronto_shelters_clean |> 
 mutate(occupancy_month = month(occupancy_date, 
                                label = TRUE,
                                abbr = TRUE)) |>
  arrange(month(occupancy_date)) |>
  drop_na(occupied_rooms) |> 
  summarise(avgs_of_rooms_occupied = mean(occupied_rooms),
             .by = occupancy_month) |>
  ggplot(aes(x= occupancy_month, y = avgs_of_rooms_occupied)) + geom_point()
```

# The occupancy of shelter's room in 2022 Toronto by different city, sector, and programs

The first graph @fig-across_cities is a scatter plot that depicts the distribution of occupied rooms by several sectors, such as families, males, mixed adults, women, and youth, across several places, including Etobicoke, North York, Scarborough, Toronto proper, and Vaughan. Each dot shows the amount of rooms inhabited by a sector within a certain city. The most notable finding here is that Toronto has the highest occupancy rate across all categories, notably for families and mixed adults, showing a concentrated demand for shelters in the city's core. 
The second graph @fig-programs presents a stacked bar chart that compares emergency and transitional shelter occupancy in the same areas. Emergency shelters have substantially greater occupancy rates, particularly in Toronto, where the bar exceeds 30,000. This implies a strong reliance on emergency shelters, which are normally used for short-term stays, implying that there may be an immediate and pressing need for shelter that cannot be provided by longer-term solutions. Transitional shelters, which provide a more stable but not permanent housing alternative, have substantially lower usage, with Toronto once again leading in occupancy.

```{r}
#| label: fig-across_cities
#| fig-cap: Occupancy of Shelter Rooms by Sector Across Toronto Cities
#| echo: false
#| warning: false
#| message: false
toronto_shelters_clean |>
  drop_na(occupied_rooms, sector, location_city) |>
  ggplot(aes(x = location_city, y = occupied_rooms, color = sector)) +
  geom_point() +
  theme_minimal() +
  labs(x = "location_city", y = "occupied_rooms", color = "Sector") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-programs
#| fig-cap: "The shelter rooms usage based on different sector and services program in Toronto."
#| echo: false
#| warning: false
#| message: false
toronto_shelters_clean |>
  drop_na(location_city) |>
  ggplot(mapping = aes(x = location_city, fill = sector)) +
  geom_bar() +
  theme_linedraw() +
  labs(x = "location_city",y = "num_occupied_rooms",fill = "sector") +
  facet_wrap(vars(program_model), scales = "free") +
  guides(x = guide_axis(angle = 60)) +
  theme(legend.position = "bottom") + 
  scale_fill_viridis_d(option = "magma")
```

# Measurements
The data measures demonstrate the systemic influence of COVID-19 and other crises on shelter demands when paired with the contextual information given in the introduction and abstract. The high occupancy rates, particularly in emergency shelters, indicate a severely overburdened shelter system despite efforts like the establishment of emergency shelters and the shift to permanent housing. This thorough data analysis confirms the necessity for well-informed policy responses that improve the support given to different demographic groups and the system's capability, especially in light of impending and continuing challenges.
Data on shelter occupancy in the Toronto area is shown in a scatter plot and bar graph. The first graph summarises total occupancy in emergency versus transitional shelters, while the second shows individual occupancy by sector (families, men, mixed adults, women, and youth) across several cities. The scatter plot's data points give an overview of demand within each demographic sector. At the same time, the graph's stacked bars show the overall occupancy levels and emphasize the need for emergency shelters, especially in central Toronto. The y-axes of both visualizations quantify the number of rooms occupied and use numerical measures of occupancy.

# Model

In our study, the study used a Bayesian analysis model to investigate occupancy patterns and service utilization in Toronto's shelter system in 2022. The model was chosen because of its ability to account for prior knowledge and uncertainty, resulting in a probabilistic approach to understanding shelter usage data. The model used a hierarchical structure to account for the different effects of factors like location, demographic groups, and time of year on shelter occupancy rates. The posterior distributions of model parameters revealed information about the relationships between these factors and shelter usage, particularly across different parts of the city and demographic groups.

However, The Bayesian model used in the analysis was effective in drawing significant conclusions about Toronto's shelter service usage trends. The study provided a multidimensional view of the shelter system by combining quantitative analysis and data visualization, as well as identifying significant trends and patterns that can inform service delivery and policy. Instead of being based on an image, the information about the priors being normally distributed with a location of 0 and a scale of 2.5 is based on a standard Bayesian statistics method for defining model priors. A normal distribution is frequently used as a weakly informative prior. This option represents a baseline assumption that, prior to data observation, the coefficients should be centered around 0, indicating no effect with a degree of uncertainty represented by the scale.

The model specified occupancy rates as a function of several covariates, including time (months), city location (e.g., Etobicoke, North York, Scarborough, Toronto, Vaughan), and shelter demographics. Specifically, we used the number of occupied rooms (number_occupied) as a response variable, with (occupancy_month) and (location_city) as predictor variables, as well as sectors to account for different demographic groups.

## Model set-up

$y_i | \mu_i \sim \text{Normal}(\mu_i, \sigma)$: Following the Bayesian model, we know that this implies that the outcome variable $y_i$, conditioned on the mean $\mu_i$, has a normal distribution with standard deviation $\sigma$. For the Toronto shelter, $y_i$ can represent the number of people who come to the shelter on day $i$, or other relevant response metric, such as the number of occupied beds.

$\mu_i = \beta_0 + \beta_1 \cdot x_i$ : This is the linear model, where $\mu_i$ represents the expected value of $y_i$. The intercept $\beta_0$ and slope $beta_1$ are multiplied by the predictor $x_i$. In the Toronto shelter, $x_i$ represents a variable of interest, it  could represent the variables such as the capacity of the shelter, the number of services offered, location factors, or the time of year.

$\beta_0, \sim \text{Normal}(0, 2.5)$ : The intercept term $\beta_0$ is assumed to be centred around 0 with a standard deviation of 2.5. This means that you do not expect a strong relationship between the predictor variable $x_i$ and the outcome $y_i$ before analyzing the data, but you are open to the possibility of discovering one.This expresses modest prior knowledge about shelter occupancy rates, allowing the data to significantly influence the posterior distribution $\beta_0$.

$\beta_1 \sim \text{Normal}(0, 2.5)$ : In a comparable way, this indicates the previous consensus regarding the slope β1 was that it has a standard deviation of 2.5 and is centered around 0. This implies that, prior to data analysis, you do not expect a strong relationship between the predictor variable $x_i$ and the outcome of $y_i$.

$\sigma \sim \text{Exponential}(1)$ : This implies that the prior belief for the standard deviation of the outcome variable's distribution is an exponential distribution with a rate parameter of one. This suggests lower σ values, indicating less variability in daily shelter use. However, you are open to learning from the data if there is more variability than expected.This distribution is commonly used for scale parameters like σ in Bayesian models because it is constrained to positive values and can represent a wide range of variability.

Furthermore, the Bayesian model is used in our study to examine occupancy rates in Toronto's shelter system, allowing for a thorough understanding of the uncertainties and variables associated with shelter use. This method generates preliminary estimates, which can be refined as more data becomes available, resulting in a probabilistic forecast. The model's predictive power stems from its ability to differentiate between the effects of various factors on the number of people seeking shelter. By assigning priors to model parameters and tracking their evolution through posterior distributions, the model learns about the relative weight of each factor, allowing it to identify which signals are genuine and worth paying attention to.
The Bayesian framework allows for a dynamic dialogue between data and theory, challenging and refining our understanding of the relationship between socioeconomic factors and shelter demand. This dialogue lays the groundwork for developing targeted policies that are both responsive to immediate needs and proactive in promoting long-term solutions to homelessness. The model's findings serve as an empirical foundation for resource allocation discussions, allowing policymakers to prioritize funding, tailor services to demographic needs, and design interventions that directly address identified causes of shelter overuse.

\begin{align}
y_i|\mu_i &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \c x_i \\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\beta_1 &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}(1)
\end{align}

### Model justification

The Bayesian model was used to analyzing in Toronto's shelter occupancy rates since it is reliable, adaptable, and possesses a lot of information. The model analyzes what it currently knows and adds it to the data it has seen to make an accurate assessment that is based on both past events and present conditions. It finds possible correlations in nested data, which lets you get an accurate understanding of how individual and group-level covariates affect occupancy rates. Thus, the Bayesian approach measures uncertainty by giving a full probabilistic account of all unknowns.

However, the model are also flexible, since it can be updated all the time as new data comes in. This makes it a dynamic tool for ongoing policy change and assessment. The Bayesian model also lets us examine at the complicated connections and interactions between covariates like demographic variables, economic factors, and neighbourhood characteristics. It Figuring out these connections can help you see deeper patterns and trends that other modelling methods might miss. Lastly, the Bayesian model was picked because it is reliable, adaptable, and gives a lot of information. This makes it an important tool for understanding the Toronto shelter system and making it work better and have a bigger effect.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

toronto_shelters_data_first_model_rstanarm <-
  readRDS(file = here::here("models/shelters_data_model_one_rstanarm.rds"))
```

```{r}
#| label: fig-models
#| fig-cap: "The relationship between the occupancy rate and actual room occupied."  
#| echo: false
#| warning: false
#| message: false
draws <- 1000

occupancy_rate_rooms <- runif(n = draws * 16, min = 0.5, max = 1.5)

toronto_shelters_clean <-
  tibble(
    sigma = rep(rexp(n = draws, rate = 1), times = 16),
    beta_0 = rep(rnorm(n = draws, mean = 0, sd = 2.5), times = 16),
    beta_1 = rep(rnorm(n = draws, mean = 8, sd = 2.5), times = 16),
    five_km_time = rep(15:30, each = draws),
    occupancy_rate_rooms = occupancy_rate_rooms, 
    mu = beta_0 + beta_1 * occupancy_rate_rooms
  ) |>
  rowwise() |>
  mutate(
    capacity_actual_room = rnorm(n = 1, mean = mu, sd = sigma)
  )

ggplot(toronto_shelters_clean, aes(x = occupancy_rate_rooms, y = capacity_actual_room)) +
  geom_point(alpha = 0.2, color = "blue") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +  theme_minimal() + labs(title = "Analysis of Shelter Capacity vs. Occupancy Rate",
    x = "Occupancy Rate of Rooms",
    y = "Actual Room Capacity"
  )
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of Toronto's shelters usage based on occupancy rate  and service user count"
#| warning: false

modelsummary(toronto_shelters_data_first_model_rstanarm)
```

# Discussion

## Findings {#sec-first-point}

This study aimed to analyze the trends in shelter usage across Toronto in 2022, focusing on average usage rates of beds and rooms, the impact of different shelter program models, and the capacity needs across various demographics and city regions. The findings reveal significant insights into the dynamics of shelter usage and the pressing needs within the community, reflecting broader social issues.

The analysis indicates a clear seasonal trend in shelter usage, with occupancy rates peaking during the colder months. This pattern underscores the critical role of weather in influencing the demand for shelter services, aligning with previous research highlighting the vulnerability of homeless populations to adverse weather conditions (Reference). Furthermore, the distinction between emergency and transitional shelter types revealed diverse needs within the homeless population, suggesting a complex landscape of shelter requirements that vary not only by season but also by individual circumstances.

## Expanding Solutions: Toronto's Comprehensive Strategy Against Homelessness

Toronto's government has been proactive in responding to the homelessness crisis, with several plans aimed at alleviating the strain on the city's shelter system.

For the winter of 2023/2024, the City of Toronto has laid out a comprehensive Winter Services Plan to support those experiencing homelessness. This plan includes the activation of Warming Centres when temperatures drop to critical lows, addition of up to 180 shelter spaces by optimizing the existing shelter sites, and the creation of 170 spaces at Warming Centres during periods of cold weather. Importantly, the city acknowledges the increasing demand for shelter and housing, highlighting the need for urgent support from federal and provincial governments to manage capacity pressures [@Toronto2023WinterServices].

Furthermore, Toronto has been actively working on creating more permanent solutions to homelessness. The city has purchased properties to be transitioned into affordable and supportive housing, with projects at 222 Spadina Ave. and 4626 Kingston Rd. already underway, providing homes for those at risk of or currently experiencing homelessness [@Toronto2023WinterPlanHomelessness]. This approach not only addresses immediate needs but also focuses on long-term solutions to prevent chronic homelessness by moving individuals from shelter systems into stable housing with necessary supports.

These efforts reflect a multifaceted approach to tackling the challenges highlighted in our paper, demonstrating the city's commitment to not just managing but resolving the issue of homelessness. The initiatives underscore the importance of collaborative efforts between different levels of government and the community to address a complex issue that affects many in Toronto. The city's plans align with our findings, suggesting that while significant steps are being taken, continued and enhanced support is necessary to meet the growing demand for shelter and affordable housing in the face of various contributing factors such as pandemic, the opioid crisis, and an influx of refugee claimants and asylum seekers.

## Implications

The implications of these findings are multifaceted. Firstly, the seasonal surge in shelter demand highlights the necessity for scalable solutions in Toronto's shelter services to accommodate fluctuating needs. This could involve dynamic resource allocation strategies that adjust in anticipation of seasonal trends. Secondly, the varied demands between emergency and transitional shelters suggest a need for a more nuanced approach to shelter provision, one that caters to the specific needs of different demographic groups.

## Limitations

This study is not without limitations. The reliance on publicly available data may introduce biases, as not all instances of shelter usage or homelessness are captured within the dataset. Additionally, the analysis does not account for the potential impact of policy changes or other external factors that may influence shelter demand. These limitations suggest caution in generalizing the findings without consideration of the broader context.

## Suggestions for Future Research

Future research should aim to address the identified limitations by incorporating more comprehensive datasets, potentially including qualitative data that captures the lived experiences of those utilizing shelter services. Furthermore, exploring the impact of policy interventions on shelter usage trends could offer valuable insights into effective strategies for managing shelter capacity and meeting the needs of vulnerable populations.

## Conclusion

The study of Toronto's shelter usage in 2022 provides critical insights into the challenges faced by homeless populations and the services designed to support them. By understanding the trends and factors influencing shelter demand, policymakers and service providers can better prepare for and address the complex needs of those seeking shelter. Continued research and innovation in this field are essential to developing effective solutions that ensure no individual is left without refuge in times of need.

\newpage

# References