### 2024-03-16 paper/datasheet_template.qmd
update for datasheet

---
title: "Datasheet for 'Beyond Capacity: Analyzing Shelter Use and Government Responses to Homelessness in Toronto'"
author: 
  - Jerry Lu (Yen-Chia Lu)
  - Sinan Ma 
  - Che-Yu Wang
thanks: "Code and data are available at: https://github.com/Sinanma/shelter_in_Toronto"
date: "`r Sys.time()`"
date-format: "12 March, 2024"
abstract: "The study examines the impact of the COVID-19 pandemic and government policies on Toronto's shelter system in 2022. It analyzes the system's response to external pressures using a dataset that includes shelter occupancy rates, service user counts, and capacity metrics. The findings show significant fluctuations in shelter demand and occupancy, highlighting both the system's resilience and vulnerabilities. The study also offers policymakers valuable insights into how to strengthen support for homeless populations in the face of future challenges. Despite temporary emergency shelters and a plan to transition to permanent affordable housing, the shelter system was under pressure from the pandemic, opioid crisis, and an increase in refugee claimants. The analysis offers useful insights for policymakers looking to improve support for homeless people."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The purpose of the dataset creation was to investigate how Toronto's shelter system might be affected in 2022 by the COVID-19 pandemic and governmental initiatives. It sought to close the knowledge gap on how these variables affected the demand for and occupancy of shelters and to offer suggestions for enhancing services for the homeless. The study investigated service patterns and system capacity issues using a Bayesian technique and visual data analysis.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was produced by OpenDataToronto, a City of Toronto effort that aims to provide public access to city data freely available for a range of purposes, such as analysis, research, and application development. This organisation is dedicated to improving openness, encouraging citizen involvement, and stimulating creativity by providing important city data.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the dataset by OpenDataToronto was funded by the City of Toronto as part of its operational budget for public services. 
4. *Any other comments?*
    - Nope

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The shelter system in Toronto in 2022 dataset includes capacity parameters, service user counts, and shelter occupancy rates. These components offer an understanding of how the shelter network responds to shocks like the COVID-19 epidemic by reflecting the day-to-day activities, capacity, and demand of the system. The dataset provides a comprehensive understanding of the relationships between the city's homeless population and the resources available to support them through a combination of quantitative data on shelter availability and utilisation.
2. *How many instances are there in total (of each type, if appropriate)?*
	- Around 27 in total. 
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset are contain all possible instances for a larger set. 
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- The dataset was included a lots of Raw data. For example, CAPACITY_ACTUAL_ROOMS, LOCATION_CITY, OCCUPANCY_RATE_ROOMS, etc. 
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, the data are target associated with the real usage of shelters in Toronto. 
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Yes, the reseaon might be the privacy concerns, or the data collection limited. 
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, the relationships between specific instances in the dataset are made clear. For instance, a city's location, sector, and occupancy rate can all influence the various sexual and usage patterns found there.  
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No, we thinking the data from opendatatoronnto are very professional. 
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No, we thinking the data from opendatatoronnto are very professional. 
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- While the OpenDataToronto dataset on Toronto's shelter system in 2022 is mainly self-contained, it may make use of other sources for further context or analysis, including academic articles, government reports, and other datasets. These external resources might not be included in official archive versions that contain the external resources as they were at the time the dataset was developed, and there are no guarantees that they will stay consistent over time. Depending on where they came from, external resources may have their own limitations, such as fees or licence requirements. One should consult the sources directly or the dataset's description for information on access and any related limitations.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- OpenDataToronto's dataset on Toronto's shelter system in 2022 is unlikely to contain any confidential information. It is intended for public use while adhering to privacy rules, which means that any personally identifying information is anonymized or aggregated to safeguard individual privacy. This approach assures that the dataset is suitable for study and analysis while maintaining personal anonymity.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- The OpenDataToronto dataset about Toronto's shelter system in 2022 is unlikely to contain anything that is immediately harmful, insulting, threatening, or anxiety-inducing, as it focuses on statistical statistics on shelter use and capacity. It is intended for objective research and policy-making, yet the subject of homelessness may elicit strong emotional responses due to the systemic challenges it raises.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The OpenDataToronto dataset for Toronto's shelter system in 2022 defines sub-populations based on gender, city, programs, usage and other variables, and displays their distribution as counts or percentages. This method aids in recognising the different requirements of the homeless population and developing focused responses.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- The OpenDataToronto dataset about Toronto's shelter system in 2022 is intended to ensure that no people can be identified, either directly or indirectly. It uses data anonymization and aggregation to protect privacy and comply with data protection rules, ensuring that no personal identifiers or sensitive information are included.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The OpenDataToronto dataset on Toronto's shelter system in 2022 may include demographic information such as age, gender, and perhaps race or ethnic origins, in order to analyse service utilisation and needs. However, it is unlikely to include sensitive information such as sexual orientation, religious beliefs, political ideas, financial, health, biometric, or genetic data, official identification, or criminal past. Any contained data is handled using privacy safeguards to prevent individual identification and to ensure compliance with data protection legislation.
16. *Any other comments?*
	- No

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The OpenDataToronto dataset about Toronto's shelter system in 2022 is mostly made up of data gathered from administrative records and daily operations at shelters. It includes measures that can be seen directly, such as the number of people using the shelter and its capacity. People who use services are required to give demographic information as part of standard intake processes. Validation methods described in research techniques would be used on any derived or inferred data to make sure it is correct and reliable. Because the dataset is used to plan and make policies for public services, keeping its quality and safety is very important.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The OpenDataToronto dataset on Toronto's shelter system was put together by shelter staff entering data by hand and using software to handle the data. Validation includes making sure the data is correct and doing regular audits while following strong quality and moral standards to get accurate data.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The OpenDataToronto dataset on Toronto's shelter system in 2022 is meant to be complete, with all the important data included. It is not just a small part of a bigger set. To correctly show how the system works, it has detailed information on shelter occupancy, capacity, and user demographics. This dataset doesn't usually use sampling methods because it's meant to give a full picture that can help with policy and service planning.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- People who work for the city, like shelter staff and data analysts, collected the OpenDataToronto dataset on the shelter system. They were paid for their work as part of their normal salaries. 
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data for the OpenDataToronto dataset on Toronto's shelter system was gathered all through 2022, which is the same time frame that the copies of the dataset were created. This makes sure that the occupancy rates, service user counts, and capacity data for the shelters accurately show how they were being used and how they were working during that time.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Since the OpenDataToronto dataset is made up of administrative data that is normally collected by the city, it's possible that it didn't go through a separate ethical review process by an institutional review board. 
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- We collect the data from the websites. 
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- The question notified about the data collection should be necessary. 
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- The people's consent should be obtained before any data is collected. For instance, those who are required to gather data will question those who have already been gathered regarding privacy. 
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Indeed, those who gave their consent have the option to withdraw it at a later time. 
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- We don't. 
12. *Any other comments?*
	- NO

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References